Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-12-18 paper/manuscript.tex
Updates from Overleaf

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2,12]{Maximilian Kuhn}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[5]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[2]{Jenke Scheen}
\author[6*]{Michael Shirts}
\author[11]{Gary Tresadern}
\author[10]{Huafeng Xu}

@@

\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy (AFE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqeuous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one type to another or can be removed or inserted by decoupling or coupling with their environment. Fig.~\ref{fig:fig_what_is_alchemy} illustrates common changes of free energies that may be difficult to compute with conventional computational methods, but are more readily tractable with AFE methods. In these types of simulations so-called \textit{alchemical states} permit converged calculations for large changes, such as the entire removal of a ligand from a binding site, as seen in fig.~\ref{fig:fig_what_is_alchemy} (E). This type of simulation is known as an absolute AFE calculation. Other typical problems tacked with AFE calculations can be as simple as conformational changes of a molecule, that exhibit a high barrier between these (fig.~\ref{fig:fig_what_is_alchemy} (A)), computation of partition coefficients, such as $\log P$ or $\log D$ (fig.~\ref{fig:fig_what_is_alchemy} (B))~\cite{rustenburg2016measuring, bosisio2016blinded}, and membrane insertions (fig.~\ref{fig:fig_what_is_alchemy} (C))~\cite{}. Then there are common types of AFE that estimate free energies of binding: the ligand stays the same, but a protein residue is alchemically mutated and the effect on the binding affinity is probed (fig.~\ref{fig:fig_what_is_alchemy} (D))\cite{hauser2018predicting,aldeghi2018accurate}, changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}, absolute binding free energies (fig.~\ref{fig:fig_what_is_alchemy} (E))~\cite{}, and relative free energies of binding where the difference of binding free energy between to similar ligands is evaluated (fig.~\ref{fig:fig_what_is_alchemy} (F))~\cite{}.

In order to obtain a free energy from these AFE simulations, historically two common approaches were used for the analysis: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or Bennet's acceptance ratio~\cite{bennett1976efficient} and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. This foundation work, combined with the methodological, technological, and hardware improvements of the last 5-10 years have led to an explosion of interest. As the field of molecular simulation moves towards millisecond timescales accurate AFE calculations on even more challenging problems will be in scope. In the meantime, today's user may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This best practice guide aims to serve as a go-to reference for current recommendations and tips for users of all experience.

@@

As some of the theoretical background can seem daunting, we provide an essentials guide to the theory behind alchemical free energy calculations in Section~\ref{sec:theory}.
We will cover topics that are key to the preparation~\ref{sec:step1}, choice and use of correct protocols~\ref{sec:step1}, and finally the best practices that should be used for the analysis of AFE calculations~\ref{sec:step3}. 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).

@@

\begin{tcolorbox}[title=Common terminology, colback=green!10!white]
    ligand - small molecular binding to a larger host molecule such as a protein. 
    lead optimization
    thermodynamic cycle
\end{tcolorbox}

@@

When starting an alchemical free energy project you need to determine if an alchemical free energy calculation is the appropriate strategy by initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are alchemical free energy calculations?}

@@

\todo[inline,color={green!20}]{ASJSM: Discuss implications of the SAMPL6 SAMPLing paper~\cite{rizzi2019sampl6} on reproducibility between simulation packages and potential sensitivity to other calculation details. @jmichel, @dmobley}
Finite computing resources necessarily limit the generated number of uncorrelated samples of potential energy surfaces, and therefore alchemical free energy calculations only give free energy estimates to within given precision. An important consideration is how reproducible alchemical free energy calculation are in practice. In simple cases such as absolute hydration free energies of small organic molecules, or relative hydration free energy calculations between structurally similar organic molecules it should be possible to obtain highly precise estimates with a given software package (with a sample standard deviation under 0.01 kcal/mol). For more complex use cases such as protein-ligand binding free energies the repeatability is often substantially worse. A good practice is to perform two or three runs of the same perturbation to assess repeatability with a given protocol. The sample standard deviation will give a crude estimate of the reliability of the estimates, and whether the precision is sufficient for the problem at hand. Where this is practical, a more stringent test is to use different input coordinates for each repeat run.Note that these issues concern calculations carried out with a single software package. The issue of reproducibility of free energy calculations across different simulation packages has attracted attention recently. Greater variability is expected due to methodological differences such as integrators, thermostats, barostats, treatment of long-range electrostatics. For absolute and relative hydration free energies of small organic molecules a variability of ca. 0.2 kcal/mol between popular simulation packages has been reported.(REF Reproducibility of hydration free energies Loefler et al.  JCTC 2018) In recent SAMPL6 SAMPLing challenge a larger variability of 0.3 to 1.0 kcal/mol was noted in the computed absolute binding free energies of host/guest systems even though similar input and simulation parameters were used (Ref Rizzi et al SAMPLing biorxiv). Further work is needed to ensure reproducibility of alchemical free energy calculations across implementations to guarantee that force-field development efforts lead to transferable potential energy functions.

@@

\subsection*{Is the expected accuracy of the computation sufficient?}
Additionally, the requisite level of accuracy is important---if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1--2 kcal/mol range~\cite{mobley2012perspective}, but if the number of compounds to be

synthesized is very small, this accuracy may not be enough to provide
much value.

@@

The concerns mentioned in the previous section such as understanding if the series of molecules is suitable, the simulation timescales relevant to capture the structure activity relationships (SAR), and performing up-front tests of performance are all relevant to drug discovery applications. However, without venturing too far into details of system setup, which is beyond the scope of this article, there are some additional points to consider. 

@@

The calculations aim to capture the alchemical change from one ligand to another as accurately as possible. Therefore, it is necessary to consider details of the experimental setup, such as pH. Biological assays are usually run at neutral pH but is not always the case, and protein and ligand preparation protocols should reflect this. In more detail, the formal charge, or tautomeric state of the small molecules can change within a series of analogs. The change of pKa is a common medicinal chemistry strategy to modify drug properties and would normally require explicit effort to incorporate into alchemical calculations. Further concerns about the experimental setup require understanding what protein construct is used in the bioassay. For instance, does the X-ray that is to be used for the calculations match the construct used for screening (i.e. catalytic vs full length, monomer vs dimer) (cite Perez-Benito Sci Rep 2018, 8, 4883)? Also, were certain co-factors or partner proteins required to augment activity? 

@@

As also mentioned, good performance of alchemical FE calculations requires an accurate representation of the ligand binding mode, normally solved via X-ray crystallography. If more than one structure is available, the modeler should pay attention to choose the most suitable. The quality of the structure can be a concern, and the reader is referred to work of Warren et al. for a detailed discussion of choosing optimal structures for structure-based modeling~\cite{warren2012essential}. It is also useful to study the structure activity relationship and understand the expected impact in the binding site, is side chain movement required and is there evidence of this in any alternative X-ray structures of the same protein? Normally, only one protein and water configuration is used for a series of alchemical FE calculations, so this needs to be capable of accommodating the smallest through to largest ligands in a way that allows stable and well behaved simulations. This can be a practical limitation for the alchemical changes that are feasible, a simple work around can be to separate compounds into sub-series for different calculations. If multiple structures are available there is some evidence the higher affinity complex can give better performance~\cite{perez-benito2019predicting}.

@@

In a drug discovery setting it is normal to consider dozens (or more) of ligands and it is necessary to align them in the binding site. There is no detailed study of how different alignment approaches may affect results, but the user should be aware of some practical considerations. Tools are available to compare the ligands and build the hybrid topologies that define the changes between one ligand and another~\cite{loeffler2015fesetup,hedges2019biosimspace,gapsys2015pmx}. In simple terms, providing poor alignment to these tools will make this job harder. Therefore, docking ligands with no restraints is not usually the best method for defining the ligand inputs, instead a core restrained docking can provide satisfactory outcomes, but in this case one still should pay careful attention to consistency of alignment for identical substituents. Another alternative often used is to manually edit the same core with the changing substituents. This provides assurances that coordinates for the non-perturbed structures remain identical and aromatic substituents for instance obey consistent dihedral angles. This approach is not feasible for many compounds and therefore automating this approach can be desirable. Finally, the role of water in ligand binding is now well understood and it is a crucial aspect to accurately reflect the changes in binding site solvation during ligand binding. Can crystallographic waters be retained? Do they clash with some of the larger ligands used in the FEP? If so, how will the water be incorporated for smaller ligands that require the water – consider pre-solvation methods such as GCMC steps~\cite{michel2010prediction}. Tools such as watermap or opensource equivalents (SSTmap, others) can be used to help define water structure for systems with no experimental evidence of water sites. Before launching large numbers of alchemical free energy calculations it is recommended to test the defined system using classical MD simulations and limited numbers of alchemical perturbations. Metrics such as ligand and protein RMSD and RMSF can be inspected, along with visual inspection of simulations, to ensure the system is stable and likely to be suitable for FE calculations. 
Running binding FE calculations in a drug discovery application will typically require the use of software or tools to facilitate the calculations of many compounds. The commercial implementation from Schr\"{o}dinger, FEP+, uses a graphical user interface and once the user has calculated any missing force field parameters, calculations can be prepared and submitted within minutes. The software builds the network of perturbations, connecting similar compounds based on LOMAP methodology~\cite{liu2013lead}, all inputs for the Desmond simulations are built automatically and the job can be saved or run immediately on available GPUs. The cost of this software is often prohibitive but alternative tools are becoming available to allow relatively straight forward running of alchemical FE calculations with other software~\cite{gapsys2015pmx, loeffler2015fesetup} (CITE New Merz paper, cite PMX paper Vytas and Bert de Groot, cite FESETUP, cite FEanalysis). One consideration at this point is to define the perturbations suitable for the application. For instance, if the goal is to accurately assess the relative binding energy of a small number of compounds, possibly with challenging synthesis, the map of perturbations may contain as many connections between compounds as affordable. However, when running calculations on hundreds of compounds a so called ‘star-map’ can be used that just contains one connection per compound: perturbing every compound to a central ligand, typically the crystal structure ligand. In this way the top-ranking examples can be readily identified and submitted to additional calculations in a second round. 
\subsection*{Making predictions, understanding errors}
For prospective drug discovery applications there are several other considerations including understanding likely errors and taking selection bias into account. It is crucial when proposing compounds for synthesis to have some idea of the underlying error or uncertainty in the predictions. Fortunately, there are several useful indicators, results suggest MBAR errors for perturbations greater than 0.3 kcal/mol can be indicative of inaccuracy~\cite{perez-benito2019predicting}. In addition, the hysteresis can be checked by comparing the difference for forward and reverse perturbations of the same structural change, or in the case of large maps of perturbations cycle closure hysteresis~\cite{wang2013modeling} indicates problematic perturbations involved in cycles connecting many compounds. Selection bias, as described above, can be problematic in prospective applications as there is normally an aim to focus on only a narrow range of predicted activity (i.e. the most active). Thus, correlation between prediction and subsequent experiment may lead to disappointment but corrections can be applied based on previous recommendations~\cite{abel2017critical}. 

@@

Many relative free energy planning tools have been set up to avoid changing the net charge of the systems considered, including LOMAP~\cite{} and early implementation of Schr\"{o}dinger's FEP+, though later implementations allow changes in net charge by including charge corrections.

Absolute free energy calculations can potentially avoid changing the charge of the system by making a charge perturbation of equal and opposite sign elsewhere in the system; for example, as a charged ligand is removed, a charged counterion of opposite sign could also be removed, or one of the same sign could be inserted.
This is the approach employed by the Yank free energy package~\cite{wang2013identifying}.

Charge corrections have also been explored, and are potentially a viable solution to this problem~\cite{} where artifacts introduced by finite-size effects are corrected numerically.
~\cite{chen2018accurate}). However, application of such corrections typically remains less common than the use of a co-alchemical ion.

@@

Once data has been collected from alchemical intermediates using a sampling algorithm described above, this data must be analyzed to produce an estimate of the transfer free energy (and its associated statistical uncertainty) for each edge of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to report on poor convergence.

@@

To minimize this effect, an initial portion of the simulation is often discarded to \emph{equilibration}, with the idea of removing the most heavily biased initial portion of simulation data but retaining the unbiased \emph{production} region that represents a stationary Markov chain process sampling from the desired equilibrium target distribution.
Because the simulation time required for the atypical initial sampler state to relax toward equilibrium is a property of the specific system being simulated and the specific initial conditions selected, it is simplest to collect data for the whole process and use an automated algorithm to select how much data should be discarded to equilibration in a post-processing step.

@@

@@

\paragraph{Cautions in automating equilibration detection}
For simulations that are simply not long enough to contain a large number of samples from true equilibrium either because they are very short or contain slow processes, this procedure cannot completely remove the bias.
In those cases, this approach simply selects the final portion of the the simulation, which may be contained in a single substate of conformational space, which may itself lead to biased estimates. 
This situation can be detected if the equilibration boundary $t_0$ is a significant fraction of the total simulation length $T$, with a good rule of thumb being that $T \gtrsim 20 t_0$.
If this is not possible, advanced analysis techniques that assume only local equilibrium (rather than global equilibrium) such as the TRAM estimators~\cite{mey2014xtram,wu2016multiensemble,nuske2017markov} may be more appropriate.

@@

\paragraph{Cautions and considerations}
Reliable estimation of the statistical inefficiency is difficult, and estimates will not generally be as precise (in a relative error sense) as averages.
To ensure there is sufficient data available for reliable decorrelation and estimation of free energy differences, it is recommended that the effective number of uncorrelated samples $N_{\mathrm{eff}} \ge \approx 50$.

@@

Free energy differences between two different states differing in the energy function are directly related to the
ratio of probabilities of those states.
As can be noted, the partition functions in Eq.~\ref{equation:dimensionless-free-energy-difference} are simply the total accumulated probabilites for all possible configurations of the system. Virtually all of the ways to estimate this free energy are based in converting this ratio of integrals to something that can be measures in one (or several) simulations.

\paragraph{The Zwanzing relationship (EXP)}

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought that term later grew to encompass all ways of calculating free energy differences.

>>>>

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[2*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2,12]{Maximilian Kuhn}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[5]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[2]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[11]{Gary Tresadern}
\author[10]{Huafeng Xu}

@@

\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy (AFE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqeuous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one type to another or can be removed or inserted by decoupling or coupling with their environment.

Fig.~\ref{fig:fig_what_is_alchemy} illustrates common free energy changes that may be difficult to compute with conventional computational methods, but are more readily tractable with AFE methods. In these types of simulations so-called \textit{alchemical states} permit converged calculations for large changes, such as the entire removal of a ligand from a binding site, as seen in fig.~\ref{fig:fig_what_is_alchemy} (E). This type of simulation is known as an absolute AFE calculation. Other typical problems tacked with AFE calculations can be as simple as conformational changes of a molecule, that exhibit a high barrier between these (fig.~\ref{fig:fig_what_is_alchemy} (A)), computation of partition coefficients, such as $\log P$ or $\log D$ (fig.~\ref{fig:fig_what_is_alchemy} (B))~\cite{rustenburg2016measuring, bosisio2016blinded}, and membrane insertions (fig.~\ref{fig:fig_what_is_alchemy} (C))~\cite{}. Then there are common types of AFE that estimate free energies of binding: the ligand stays the same, but a protein residue is alchemically mutated and the effect on the binding affinity is probed (fig.~\ref{fig:fig_what_is_alchemy} (D))\cite{hauser2018predicting,aldeghi2018accurate}, changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}, absolute binding free energies (fig.~\ref{fig:fig_what_is_alchemy} (E))~\cite{}, and relative free energies of binding where the difference of binding free energy between to similar ligands is evaluated (fig.~\ref{fig:fig_what_is_alchemy} (F))~\cite{}.

In order to obtain a free energy from these AFE simulations, historically two common approaches were used for the analysis: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or Bennett's acceptance ratio~\cite{bennett1976efficient} and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. This foundational work, combined with the methodological, technological, and hardware improvements of the last 5-10 years have led to an explosion of interest.

As the field of molecular simulation moves towards millisecond timescales accurate AFE calculations on even more challenging problems will be in scope. In the meantime, today's users may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This best practice guide aims to serve as a go-to reference for current recommendations and tips for users of all experience.

@@

As some of the theoretical background can seem daunting, we provide an essentials guide to the theory behind alchemical free energy calculations in Section~\ref{sec:theory}.
We will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of AFE calculations~(\ref{sec:step3}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).

@@

\begin{tcolorbox}[title=Common terminology, colback=green!10!white]
    ligand - small molecular binding to a larger host molecule such as a protein.\\ 
    lead optimization - process of designing a drug candidate from an initial lead compound, considering its affinity and various pharmacokinetic parameters.\\
    thermodynamic cycle - sequence of thermodynamic processes (like changes in temperature, volume or pressure) in which the initial and final state are equal.
\end{tcolorbox}

@@

When starting an alchemical free energy project you need to determine if an alchemical free energy calculation is the appropriate strategy by initially counting the cost of your
project: Can you even hope to tackle the problem with available resources and, if successful, will it
be worth it?
\subsection*{How accurate are alchemical free energy calculations?}

@@

\todo[inline,color={green!20}]{ASJSM: Discuss implications of the SAMPL6 SAMPLing paper~\cite{rizzi2019sampl6} on reproducibility between simulation packages and potential sensitivity to other calculation details. @jmichel, @dmobley}
Finite computing resources necessarily limit the generated number of uncorrelated samples of potential energy surfaces, and therefore alchemical free energy calculations only give free energy estimates to within a given precision. An important consideration is how reproducible alchemical free energy calculations are in practice. In simple cases such as absolute hydration free energies of small organic molecules, or relative hydration free energy calculations between structurally similar organic molecules, it should be possible to obtain highly precise estimates with a given software package (with a sample standard deviation under 0.01 kcal/mol). For more complex use cases such as protein-ligand binding free energies the repeatability is often substantially worse. A good practice is to perform two or three runs of the same perturbation to assess repeatability with a given protocol. The sample standard deviation will give a crude estimate of the reliability of the estimates, and whether the precision is sufficient for the problem at hand. Where this is practical, a more stringent test is to use different input coordinates for each repeat run. Note that these issues concern calculations carried out with a single software package. The issue of reproducibility of free energy calculations across different simulation packages has attracted attention recently. Greater variability is expected due to methodological differences such as integrators, thermostats, barostats, treatment of long-range electrostatics. For absolute and relative hydration free energies of small organic molecules a variability of ca. 0.2 kcal/mol between popular simulation packages has been reported.(REF Reproducibility of hydration free energies Loefler et al.  JCTC 2018) In the recent SAMPL6 SAMPLing challenge a larger variability of 0.3 to 1.0 kcal/mol was noted in the computed absolute binding free energies of host/guest systems even though similar input and simulation parameters were used (Ref Rizzi et al SAMPLing biorxiv). Further work is needed to ensure reproducibility of alchemical free energy calculations across implementations to guarantee that force-field development efforts lead to transferable potential energy functions.

@@

\subsection*{Is the expected accuracy of the computation sufficient?}
The requisite level of accuracy is another important consideration ---if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracy in the 1--2 kcal/mol range~\cite{mobley2012perspective}, but if the number of compounds to be synthesized is very small, this accuracy may not be enough to provide
much value.

@@

It is expected that AFE methods can provide a valuable tool for drug discovery, capable of outperforming alternative approaches. (CITE Kuhn et al J. Med. Chem. 2017, 60, 6, 2485-2497). It has been stressed that successful application in industry requires more knowledge of the "domain of applicability", in other words, a greater understanding of the success and failure~\cite{sherborne2016collaborating}, which in turn requires robust protocols for preparing, submitting and analysing AFE calculations. In this regard, the issues mentioned in the previous section such as understanding the suitability and timescales to capture the structure activity relationships (SAR), and performing up-front tests of performance are all relevant to drug discovery applications. Without venturing too far into details of system setup, which is beyond the scope of this article, there are some additional points to consider. 

@@

The calculations aim to capture the alchemical change from one ligand to another as accurately as possible. Therefore, it is necessary to consider details of the experimental setup, such as pH. Biological assays are usually run at neutral pH but this is not always the case, and protein and ligand preparation protocols should reflect this. In more detail, the formal charge, or tautomeric state of the small molecules can change within a series of analogs. The change of pKa is a common medicinal chemistry strategy to modify drug properties and would normally require explicit effort to incorporate into alchemical calculations. Further concerns about the experimental setup require understanding what protein construct is used in the bioassay. For instance, does the X-ray that is to be used for the calculations match the construct used for screening (i.e. catalytic vs full length, monomer vs dimer) (cite Perez-Benito Sci Rep 2018, 8, 4883)? Also, were certain co-factors or partner proteins required in the bioassay? 

@@

As also mentioned, good performance of alchemical FE calculations requires an accurate representation of the ligand binding mode, normally solved via X-ray crystallography. If more than one structure is available, the modeler should pay attention to choose the most suitable. The quality of the structure can be a concern, and the reader is referred to work of Warren et al. for a detailed discussion of choosing optimal structures for structure-based modeling~\cite{warren2012essential}. It is also useful to study the structure activity relationship and understand the expected impact in the binding site, is side chain movement required and is there evidence of this in any alternative X-ray structures of the same protein? Normally, only one protein and water configuration is used for a series of alchemical FE calculations, so this needs to be capable of accommodating the smallest through to largest ligands in a way that allows stable and well behaved simulations. This can be a practical limitation for the alchemical changes that are feasible, a simple work-around can be to separate compounds into sub-series for different calculations. If multiple structures are available there is some evidence the higher affinity complex can give better performance~\cite{perez-benito2019predicting}.

@@

In a drug discovery setting it is normal to consider dozens (or more) of ligands and it is necessary to align them in the binding site. There is no detailed study of how different alignment approaches may affect results, but the user should be aware of some practical considerations. Tools are available to compare the ligands and build the hybrid topologies that define the changes between one ligand and another~\cite{loeffler2015fesetup,hedges2019biosimspace,gapsys2015pmx}. In simple terms, providing poor alignment to these tools will make this job harder. Therefore, docking ligands with no restraints is not usually the best method for defining the ligand inputs, instead a core restrained docking can provide satisfactory outcomes, but in this case one still should pay careful attention to consistency of alignment for identical substituents. Another alternative is to manually edit the same core with the changing substituents. This provides assurances that coordinates for the non-perturbed structures remain identical and aromatic substituents for instance obey consistent dihedral angles but is not feasible for many compounds and therefore automation can be desirable. Finally, the role of water in ligand binding is now well understood and it can be crucial to capture the changes in binding site solvation during ligand binding. Can crystallographic waters be retained? Do they clash with some of the larger ligands used in the FEP? If so, how will the water be incorporated for smaller ligands that require the water – consider pre-solvation methods such as GCMC steps~\cite{michel2010prediction}. Other tools such as watermap or opensource equivalents (SSTmap, others) can be used to define water structure for systems with no experimental evidence of water sites. Before launching large numbers of alchemical free energy calculations it is recommended to test the system using classical MD simulations and limited numbers of alchemical perturbations. Metrics such as ligand and protein RMSD and RMSF can be inspected, along with visual inspection of simulations, to ensure the system is stable and likely to be suitable for FE calculations.

Running binding FE calculations in a drug discovery application will typically require the use of software or tools to facilitate the calculations of many compounds. The commercial implementation from Schr\"{o}dinger, FEP+, uses a graphical user interface and once the user has calculated any missing force field parameters, calculations can be prepared and submitted within minutes. The software builds the network of perturbations, connecting similar compounds based on LOMAP methodology~\cite{liu2013lead}, all inputs for the Desmond simulations are built automatically and the job can be saved or run immediately on available GPUs. The cost of this software is often prohibitive but alternative tools are becoming available to allow relatively straight forward running of alchemical FE calculations with other software~\cite{gapsys2015pmx, loeffler2015fesetup} (CITE New Merz paper, cite PMX paper Vytas and Bert de Groot, cite FESETUP, cite FEanalysis). One consideration at this point is to define the perturbations suitable for the application. For instance, if the goal is to accurately assess the relative binding energy of a small number of compounds, possibly with challenging synthesis, the map of perturbations may contain as many connections between compounds as affordable. However, when running calculations on hundreds of compounds a so called ‘star-map’ can be used that just contains one connection per compound: perturbing every compound to a central ligand, typically the crystal structure ligand. In this way the top-ranking examples can be readily identified and submitted to additional calculations in a second round. 
%
\subsection*{Making predictions, understanding errors}
For prospective drug discovery applications there are several other considerations including understanding likely errors and taking selection bias into account. It is crucial when proposing compounds for synthesis to have some idea of the underlying error or uncertainty in the predictions. Fortunately, there are several useful indicators, results suggest MBAR errors for perturbations greater than 0.3 kcal/mol can be indicative of inaccuracy~\cite{perez-benito2019predicting}. In addition, the hysteresis can be checked by comparing the difference for forward and reverse perturbations of the same structural change, or in the case of large maps of perturbations cycle closure hysteresis~\cite{wang2013modeling} indicates problematic perturbations involved in cycles connecting many compounds. Selection bias, as described above, can be problematic in prospective applications as there is normally an aim to focus on only a narrow range of predicted activity (i.e. the most active). Thus, correlation between prediction and subsequent experiment may lead to disappointment but corrections can be applied based on previous recommendations~\cite{abel2017critical}. 
%

@@

Many relative free energy planning tools have been set up to avoid changing the net charge of the systems considered, including LOMAP~\cite{} and Schr\"{o}dinger's FEP+~\cite{chen2018accurate}. Absolute free energy calculations can also potentially avoid changing the charge of the system by making a charge perturbation of equal and opposite sign elsewhere in the system; for example, as a charged ligand is removed, a charged counterion of opposite sign could also be removed, or one of the same sign could be inserted. This approach is sometimes referred to as an "alchemical ion" approach to dealing with the needed charge change, and is also employed by the Yank free energy package~\cite{wang2013identifying}.

Charge corrections have also been explored, and are potentially a viable solution to this problem~\cite{} where artifacts introduced by finite-size effects are corrected numerically.
~\cite{chen2018accurate}. However, application of such corrections typically remains less common than the use of a co-alchemical ion.

@@

Once data has been collected from alchemical intermediates, it must be analyzed to produce an estimate of the free energy change (and its associated statistical uncertainty) for each leg of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to report on poor convergence.

@@

To minimize this effect, an initial portion of the simulation is often discarded to \emph{equilibration}~\cite{https://doi.org/10.33011/livecoms.1.1.5957}, with the idea of removing the most heavily biased initial portion of simulation data but retaining the unbiased \emph{production} region that represents a stationary Markov chain process sampling from the desired equilibrium target distribution.
Because the simulation time required for the atypical initial sampler state to relax toward equilibrium is a property of the specific system being simulated and the specific initial conditions selected, it is simplest to collect data for the whole process and use an automated algorithm to select how much data should be discarded to equilibration in a post-processing step.

@@

For additional discussion of working with correlated data and autocorrelation analysis, please refer to the work on Best Practices for Quantification of Uncertainty and Sampling Quality in Molecular Simulations~\cite{https://doi.org/10.33011/livecoms.1.1.5067}.

@@

\paragraph{Cautions in automating equilibration detection}
For simulations that are simply not long enough to contain a large number of samples from true equilibrium either because they are very short or contain slow processes, this procedure cannot completely remove the bias.
In such cases, this approach simply selects the final portion of the the simulation, which may be contained in a single substate of conformational space, which may itself lead to biased estimates. 
This situation can be detected if the equilibration boundary $t_0$ is a significant fraction of the total simulation length $T$, with a good rule of thumb being that $T \gtrsim 20 t_0$.
If this is not possible, advanced analysis techniques that assume only local equilibrium (rather than global equilibrium) such as the TRAM estimators~\cite{mey2014xtram,wu2016multiensemble,nuske2017markov} may be more appropriate.

@@

\paragraph{Cautions and considerations}
Reliable estimation of the statistical inefficiency is difficult, and estimates will not generally be as precise (in a relative error sense) as averages.
To ensure there is sufficient data available for reliable decorrelation and estimation of free energy differences, it is recommended that the effective number of uncorrelated samples $N_{\mathrm{eff}} \ge \approx 50$ if the BAR or MBAR estimators (discussed below) are used; the number may need to be much higher with alternate estimators.

@@

Free energy differences between two different states differing in the energy function are directly related to the
ratio of probabilities of those states.
As can be noted, the partition functions in Eq.~\ref{equation:dimensionless-free-energy-difference} are simply the total accumulated probabilites for all possible configurations of the system. Virtually all of the ways to estimate this free energy are based in converting this ratio of integrals to something that can be measured in one (or several) simulations.

\paragraph{The Zwanzig relationship (EXP)}

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought that term is sometimes used to encompass all ways of calculating free energy differences.

The free energy between two states defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - U_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta U(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
In words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be to switch to potential energy function $U_1(\vec{q})$, and exponentially average the difference to get the free energy.  Here, the exponential average consists in multiplying by -1, exponentiating, averaging, taking the logarithm, and multiplying by -1 again.  This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies of stored snapshots.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change some conditions~\cite{xx}.

Although the Zwanzig equation is formally correct (as long as the two states considered sample the same phase space volume), it has some very important numerical issues that mean that it often performs badly for standard free energy calculations, even for small molecules.~\cite{shirts2005comparison,lu2003appropriate}  One can show that if the standard deviation of the difference $\Delta U(\vec{q}) = U_1(vec{q})-U_2(\vec{q})$ over all sampled $\vec{q}$ is large (which in this case, means only several times $k_BT)$, then very few samples contribute to the average, and the answer will be both biased and extremely noisy~\cite{xx}. Essentially, the method is dominated by contributions of rare snapshots\cite{cite Jarzynski 2005 PRE paper about typical vs dominant realizations}. There are a number of applications where it this expression is nevertheless useful, but those usually involve investigating the dependence of the free energy on small changes in parameters, not changing molecules or even single atoms in molecules~\cite{xx}
